Future challenges for Research Software Engineering

### 2025-03-25 future_challenges_for_RSEng.tex
let git ignore a few files
Fix footnotes and GitHub Actions
Summarise workshop output

% Required packages
% =================
% Your \usepackage commands go here.

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\texorpdfstring{\autref{1}}{},
Leyla Jael Castro\texorpdfstring{\autref{1}}{},
Gerasimos Chourdakis\texorpdfstring{\autref{1}}{},
Simon Christ\texorpdfstring{\autref{1}}{},
Jeremy Cohen\texorpdfstring{\autref{1}}{},
Jean-Noël Grad\texorpdfstring{\autref{1}}{},
Magnus Hagdorn\texorpdfstring{\autref{1}}{},
Toby Hodges\texorpdfstring{\autref{1}}{},
Jan Linxweiler\texorpdfstring{\autref{1}}{},
Frank Löffler\texorpdfstring{\autref{1}}{},
Michele Martone\texorpdfstring{\autref{1}}{},
Jan Philipp Thiele\texorpdfstring{\autref{1}}{},
Nick del Grosso\texorpdfstring{\autref{1}}{}
} % Authors and references to addresses
\institute{Fantasy University\autlabel{1}} % Institutes with labels
\abstract{Admittedly, it’s still only 2025 and we are actually stillworking on establishing Research Software Engineering
as a new domain with the hope of ensuring that it gets the recognition it deserves.
And while it is certainly too early to already herald the end of RSEng,
we should acknowledge that the future is in constant flux and
it is worthwhile to discuss how RSEng will/could change in response to these upcoming challenges in an open discussion.} % Abstract of the article
\keywords{Keywords go here.} % Keywords for the article

@@

\subsection{Definition and Evolution of RSEs}
How will the Definition of an RSE evolve in ten years?
What will not change in the essence of an RSE ten years from now?
Here we consider how the topics of RSEng will change in the face of the digitalization of society.
Also here: What will our topics be after version control doesn’t need to be taught anymore? Digitalisation will also feature more heavily in the domain curricula.

>>>>

% Required packages
% =================
% Your \usepackage commands go here.
\usepackage{fancybox,framed}

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\texorpdfstring{\autref{1}}{},
Leyla Jael Castro\texorpdfstring{\autref{1}}{},
Gerasimos Chourdakis\texorpdfstring{\autref{1}}{},
Simon Christ\texorpdfstring{\autref{1}}{},
Jeremy Cohen\texorpdfstring{\autref{1}}{},
Jean-Noël Grad\texorpdfstring{\autref{1}}{},
Magnus Hagdorn\texorpdfstring{\autref{1}}{},
Toby Hodges\texorpdfstring{\autref{1}}{},
Jan Linxweiler\texorpdfstring{\autref{1}}{},
Frank Löffler\texorpdfstring{\autref{1}}{},
Michele Martone\texorpdfstring{\autref{1}}{},
Jan Philipp Thiele\texorpdfstring{\autref{1}}{},
Nick del Grosso\texorpdfstring{\autref{1}}{}
} % Authors and references to addresses
\institute{\autlabel{1} Fantasy University} % Institutes with labels
\abstract{Admittedly, it’s still only 2025 and we are actually stillworking on establishing Research Software Engineering
as a new domain with the hope of ensuring that it gets the recognition it deserves.
And while it is certainly too early to already herald the end of RSEng,
we should acknowledge that the future is in constant flux and
it is worthwhile to discuss how RSEng will/could change in response to these upcoming challenges in an open discussion.} % Abstract of the article
\keywords{Keywords go here.} % Keywords for the article

@@

\subsection{Definition and Evolution of RSEs}
\begin{framed}
How will the definition of an RSE evolve in ten years?

What will not change in the essence of an RSE ten years from now?
\end{framed}
% Here we consider how the topics of RSEng will change in the face of the digitalization of society.
% Also here: What will our topics be after version control doesn't need to be taught anymore?
% Digitalisation will also feature more heavily in the domain curricula.

The digital skills of domain scientists will increase,
and software will become easier to interact with,
reducing reliance on RSEs to train newcomers on the most basic tasks
(JN: for example, versioning and archiving?).
User training sessions might focus on more advanced topics
that cater to a smaller audience, or involve mentoring more frequently.

Regarding the definition of RSEs, we will still have a spectrum of competencies,
from full professional RSEs to domain scientists who develop research software.
While the RSE core competencies\cite{Goth2024} might remain unchanged,
there might be a shift in the responsibilities of RSEs,
in the form of a partial reallocation of the training budget
to the other core competencies.

What will \emph{not} change over that timeframe is the RSE's ability to remain
flexible and embrace change. Keep an eye open and pick up tools and skills
as needed as we go along.
Communication and collaboration skills will remain essential to engage
with domain scientists and help them find technical solutions.
